---
title: "Nocc"
author: "kevin"
date: "January 2, 2021"
output: html_document
---

#SYNOPSIS

This document is result of the second project proposed in the “Reproducible Research” course by Johns Hopkins University, available at coursera. The goal of this project is to explore the U.S. NAtional Oceanic and Atmospheric Administration’s (NOAA) storm database and answer some basic questions about severe weather events

#Data Processing
```{r}

stormdata<- read.csv("repdata_data_StormData.csv",header=TRUE,sep = ",")
```
```{r}
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)
```
#Analysis
```{r}
kable(stormdata %>% 
    group_by(EVTYPE) %>% 
    summarise(total_injuries = sum(INJURIES)) %>% 
    arrange(-total_injuries) %>% 
    mutate(ranking = row_number()) %>% 
    select(ranking, everything()) %>% 
    head(., 10)
) 
```
```{r}
kable(stormdata %>% 
    group_by(EVTYPE) %>% 
    summarise(total_fatalities = sum(FATALITIES)) %>% 
    arrange(-total_fatalities) %>% 
    mutate(ranking = row_number()) %>% 
    select(ranking, everything()) %>%   
    head(., 10)
) 
```
As we can see, the ranking for the total injuries is different from the total fatalities. Since we are only interested, we will create the variable “total_health_damage” with the sum of the injuries and fatalities and get the top 10 most harmful event types:
```{r}
HarmfulEvents <- stormdata %>% 
    group_by(EVTYPE) %>% 
    summarise(total_health_damage = sum(sum(FATALITIES), sum(INJURIES))) %>% 
    arrange(-total_health_damage) %>% 
    mutate(ranking = row_number()) %>% 
    select(ranking, everything()) %>%   
    head(., 10)

HarmfulEvents
```

    Across the United States, which types of events have the greatest economic consequences?

The data provides two types of economic impact, namely property damage (PROPDMG) and crop damage (CROPDMG). The actual damage in $USD is indicated by PROPDMGEXP and CROPDMGEXP parameters. PROPDMGEXP and CROPDMGEXP can be interpreted as the following:-

    H, h -> hundreds = x100
    K, K -> kilos = x1,000
    M, m -> millions = x1,000,000
    B,b -> billions = x1,000,000,000
    The rest will be interpreted as 1

Let’s create a variable called “total_damage” and take a look at the top 10 event type with the most economic impact:
```{r}
economicImpact <- stormdata %>% 
  mutate(prop_multiplier = case_when(
                              PROPDMGEXP %in% c('h','H') ~ 100,
                              PROPDMGEXP %in% c('k','K') ~ 1000,
                              PROPDMGEXP %in% c('m','M') ~ 100000,
                              PROPDMGEXP %in% c('b','B') ~ 1000000000,
                              T ~ 1
                              ),
         crop_multiplier = case_when(
                              CROPDMGEXP %in% c('h','H') ~ 100,
                              CROPDMGEXP %in% c('k','K') ~ 1000,
                              CROPDMGEXP %in% c('m','M') ~ 100000,
                              CROPDMGEXP %in% c('b','B') ~ 1000000000,
                              T ~ 1
                              )
         ) %>% 
  group_by(EVTYPE) %>% 
  summarise(total_damage = sum(sum(PROPDMG * prop_multiplier), sum(sum(CROPDMG * crop_multiplier)))) %>% 
  arrange(-total_damage) %>% 
  mutate(ranking = row_number()) %>% 
  select(ranking, everything()) %>%   
  head(., 10) 

economicImpact
```
#Results
```{r}
HarmfulEvents %>% 
  mutate(total_health_damage = round(total_health_damage / 1000),2) %>% 
  ggplot(., aes(x = reorder(EVTYPE, total_health_damage), y = total_health_damage))+
  geom_bar(stat = 'identity')+
  coord_flip()+
  xlab("Event Type") + 
  ylab("Total number of health impact") +
  ggtitle("Weather event types impact on public health (in thousands)") +
  theme(plot.title = element_text(hjust = 0.5))+
  geom_text(aes(label = total_health_damage), position = position_dodge(width = 1), hjust = -0.3)
```
```{r}
economicImpact %>% 
  mutate(total_damage = round(total_damage / 1000000000),2) %>% 
  ggplot(., aes(x = reorder(EVTYPE, total_damage), y = total_damage))+
  geom_bar(stat = 'identity')+
  coord_flip()+
  xlab("Event Type") + 
  ylab("Total number of health impact") +
  ggtitle("Weather event types impact on public health (in billions)") +
  theme(plot.title = element_text(hjust = 0.5))+
  geom_text(aes(label = total_damage), position = position_dodge(width = 0.5), hjust = -0.05)
```